---
title: "TSA Final project"
author: "Afet Ibadova"
date: "07/06/2023"
output:
  html_document:
    theme: spacelab
    highlight: tango
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
  pdf_document:
    toc: yes
subtitle: "Estimating Value-at-Risk of a Portfolio with GARCH-Family models"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE, 
                      cache   = TRUE,
                      message = FALSE, 
                      warning = FALSE)
options(scipen = 10)
```

# Introduction

## Packages and data

Lets load all needed packages

```{r cache = F}
library(dplyr)
library(tidyverse)
library(xts)
library(zoo)
library(fBasics)  # basicStats()
library(tseries)  # jarque.bera.test()
library(FinTS)    # ArchTest()
library(car)      # durbinWatsonTest()
library(rmgarch)
library(fGarch) # e.g. garchFit()
library(quantmod)
library(rvest)
library(lubridate)
library(graphics)
library(tictoc)

source("functions/compare_ICs_GARCH.R")
source("functions/compare_ICs_ugarchfit.R")
```

# Prepraing the data

## Fetching Data

```{r}
start.date <- "2000-01-01"
stop.date  <- "2023-06-07"


# Fetch data for the S&P 500
SP500 <- getSymbols("^GSPC", src = "yahoo", 
                    from = start.date, to = stop.date, 
                    auto.assign = FALSE)

# Fetch data for DAX from Yahoo Finance
DAX <- getSymbols("^GDAXI", 
                  from = start.date, to = stop.date, 
                  auto.assign = FALSE)

# Read the CSV files for Wig20, kospi200, nikkei225
wig20 <- read.csv("wig20_d.csv",
                     header = TRUE,
                     sep = ",",
                     dec = ".",
                     stringsAsFactors = F)

kospi200 <- read.csv("^kospi_d.csv",
                     header = TRUE,
                     sep = ",",
                     dec = ".",
                     stringsAsFactors = F)
nikkei225 <- read.csv("^nkx_d.csv",
                      header = TRUE,
                      sep = ",",
                      dec = ".",
                      stringsAsFactors = F)
## Print the data frames for checking
head(wig20)
head(kospi200)
head(nikkei225)
head(SP500)
head(DAX)

str(wig20)
str(SP500)
str(DAX)
```

## Cleaning data

```{r}
## change date as row index
wig20 <- xts(wig20[, -1], 
             order.by = as.Date(wig20$Date))
kospi200 <- xts(kospi200[, -1],
                order.by = as.Date(kospi200$Date))
nikkei225 <- xts(nikkei225[, -1], 
                 order.by = as.Date(nikkei225$Date))

head(wig20)
```

Let's select only close values:

```{r}
## select only close values for all indexes
SP500 <- SP500[, 4]
names(SP500) <- "SP500"

wig20 <- wig20[, 4]
names(wig20) <- c("WIG20")

kospi200 <- kospi200[, 4]
names(kospi200) <- c("KOSPI200")

nikkei225 <- nikkei225[, 4]
names(nikkei225) <- c("NKK225")

DAX <- DAX[, 4]
names(DAX) <- c("DAX")
```

Create portfolio with 5 index

```{r}
portfolio <- SP500
portfolio$DAX <-DAX$DAX
portfolio$WIG20 <-wig20$WIG20
portfolio$KOSPI200 <- kospi200$KOSPI200
portfolio$NKK225<- nikkei225$NKK225

head(portfolio)
```

## Missing Values

Check the missing values

```{r}
any(is.na(portfolio))
```

I have missing values, because my data start is different for different variables, that's why limit start time from 2018

```{r}
portfolio <- portfolio["2018/", ]
tail (portfolio)
head(portfolio)
```

After limiting the timeframe, I have missing values, still... I assume that these missing values exist because of non-work days. The main reason is thedifferences of public holidays across the countries. Let's fill the null values with the previous values, assuming that if it is non-work days, the price doesn't change.

```{r}
portfolio <- na.locf(portfolio)
##Check
any(is.na(portfolio))
```

I have missing values, let's check the head of portfolio

```{r}
head(portfolio)
```

There is null values at the begining of dataset, that's why na.locf function wasn't fill them. Let's omit these missing values

```{r}
portfolio <- na.omit(portfolio)
##Check
any(is.na(portfolio))

head(portfolio)
tail(portfolio)
```

All Missing values are cleaned. Let's continue 🥳

## Adding log return of index seperatedly and applying weight

```{r}
portfolio$SP500_r <- 0.2 * diff.xts(log(portfolio$SP500))

portfolio$DAX_r <- 0.2 * diff.xts(log(portfolio$DAX))

portfolio$WIG20_r <- 0.2 * diff.xts(log(portfolio$WIG20))

portfolio$KOSPI_r <- 0.2 * diff.xts(log(portfolio$KOSPI200))

portfolio$NKK225_r <- 0.2 * diff.xts(log(portfolio$NKK225))


head(portfolio)
```

Creating of all indexes log returns all together

```{r}
portfolio$PORTFOLIO_r <- rowSums(portfolio[ , c("SP500_r", "DAX_r", "KOSPI_r", "WIG20_r", "NKK225_r")])

head(portfolio)
```

I have missing values in the first row because of the diff function, let's omit the first row

```{r}
portfolio <- na.omit(portfolio)
head(portfolio)
any(is.na(portfolio))
```

# Volatility Modelling - stylized facts

```{r}
## Plot of portfolio returns

plot(portfolio$PORTFOLIO_r,
     col = "blue",
     main = "Portfolio returns")
```

Checking skewness and kurtosis

```{r}
basicStats(portfolio$PORTFOLIO_r)
```

```{r}
# Plotting Histogram of Log-returns
tibble(r = as.numeric(portfolio$PORTFOLIO_r)) %>%
  ggplot(aes(r)) +
  geom_histogram(aes(y =..density..),
                 colour = "black", 
                 fill = "light green") +
  stat_function(fun = dnorm,colour="blue",
                args = list(mean = mean(portfolio$PORTFOLIO_r), 
                            sd = sd(portfolio$PORTFOLIO_r))) +
  theme_light() + 
  labs(
    title = paste("Density of the Portfolio log-returns"), 
    y = "", x = "",
    caption = "source: own calculations"
  )
```

Jarque-Bera statistic:

```{r}
jarque.bera.test(portfolio$PORTFOLIO_r)
```

**Conclusion for this part:**

1\. Skewness is negative: -0.825809

2\. Kurtosis is 13.232398, it means strong excess kurtosis

3\. The plot is leptokurtosis, "heavy tails" and higher peak in mean

4\. According to the JB test, we reject the null hypothesis about normality. The null hypothesis is a joint hypothesis of the skewness being zero and the excess kurtosis being zero.

# ARCH Test

Let's plot the ACF function of log-returns:

```{r}
acf(portfolio$PORTFOLIO_r, 
    lag.max = 36, 
    na.action = na.pass,
    ylim = c(-0.1,0.1), # we rescale the vertical axis
    col = "darkgreen", 
    lwd = 7, 
    main = "ACF of log-returns of Portfolio")
```

As seen, values of the ACF function indicate some autoregressive/MA relations among returns which can be used to build an ARIMA model.

Now, let's also see the ACF values for for the **squared** log-returns:

```{r}
acf(portfolio$PORTFOLIO_r ^ 2, 
    lag.max = 36, 
    na.action = na.pass,
    ylim = c(0,0.5), # we rescale the vertical axis
    col = "blue", 
    lwd = 7, 
    main = "ACF of SQUARED log-returns of Portfolio")
```

Let's verify existence of ARCH effects among log-returns.

The ARCH test is based on the autocorrelation of **squared** returns.

```{r}
ArchTest(portfolio$PORTFOLIO_r,  # here we use a vector of returns as input
         lags = 5) # and maximum order of ARCH effect
```

The null hypothesis about lack of ARCH effects strongly rejected!

*Note:* I check this test for different lags, for example, 5,10,15. All the lags indicate same result.

```{r}
durbinWatsonTest(lm(portfolio$PORTFOLIO_r^2 ~ 1),
                 max.lag = 5)
```

The null hypothesis for durbinWatsonTest about no autocorrelation in squared log returns is rejected.

# Modelling

Now, let's try to find the most attractive GARCH(q, p) model.

## ARCH vs. GARCH models

### ARCH(1)

```{r}
k.arch1 <- garchFit(formula = ~ garch(1, 0), # GARCH(q, p)
                    # formula describing the mean and variance equation 
                    # of the ARMA-GARCH - we assume that returns
                    # follow an ARCH(1) process
                    data = portfolio$PORTFOLIO_r,
                    # conditional distribution of errors
                    cond.dist = "norm", 
                    # if we don't want to see the history of iterations
                    trace = FALSE) 

# Summary of results and some diagnostic tests:

summary(k.arch1)
plot(k.arch1, which = 10)
plot(k.arch1, which = 11)
plot(k.arch1, which = 2)

```

The parameters are denoted by the following symbols:

-   mu - constant term in the mean equation
-   omega - constant term in the variance equation
-   alpha1 - arch1 parameter in the variance equation

If the intercept in the mean equation of the model above is not significant at 5% level, we can drop it.

Let's drop it.

### ARCH(1) - without mu

```{r}
k.arch1a <- garchFit(~garch(1, 0), 
                     data = portfolio$PORTFOLIO_r,
                     # remove intercept term in the mean equation
                     include.mean = FALSE,
                     # conditional distribution of errors
                     cond.dist = "norm",
                     trace = FALSE)
summary(k.arch1a)
plot(k.arch1a, which = 10)
plot(k.arch1a, which = 11)
plot(k.arch1a, which = 2)
```

-   Significant lags are observed for square residuals for lags 2-12, 14-15.
-   And the Ljung-Box test for R\^2 shows there is autocorrelation at 5% level.

Let's try ARCH(5).

### ARCH(5)

```{r}
k.arch5 <- garchFit(~garch(5, 0),
                    data = portfolio$PORTFOLIO_r,
                    include.mean = TRUE,
                    cond.dist = "norm", 
                    trace = FALSE) 
summary(k.arch5)

plot(k.arch5, which = 10)
plot(k.arch5, which = 11)
```

-   All parameters are significant.
-   The Ljung-box tests for R and R\^2 show there is still autocorrelation.
-   The lags 5,6,9,in ACF for squared residuals seem to be still significant.

### ARCH(10)

```{r}
k.arch10 <- garchFit(~garch(10, 0),
                    data = portfolio$PORTFOLIO_r,
                    include.mean = TRUE,
                    cond.dist = "norm", 
                    trace = FALSE) 
summary(k.arch10)

plot(k.arch10, which = 10)
plot(k.arch10, which = 11)
```

-   Again most of parameters are significant at 5% level - excluding 4,5,7,10.
-   The Ljung-box test for R\^2 shows there is no more autocorrelation between the current and past standardized squared residuals.
-   It looks like no further extension of the **conditional variance** equation is needed.

### GARCH (1,1)

```{r}
k.garch11 <- garchFit(~garch(1, 1),
                      data = portfolio$PORTFOLIO_r,
                      include.mean = TRUE,
                      cond.dist = "norm", 
                      trace = FALSE) 
summary(k.garch11)

plot(k.garch11, which = 10)
plot(k.garch11, which = 11)
```

-   All parameters are significant.
-   The Ljung-Box test for R\^2 does not show autocorrelation at 5% level between the current and past standardized squared residuals (variance).
-   The **conditional variance** equation seems to be complete, however the ACF plot of residuals indicates that we may still improve the model by adding AR components in the **mean** equation.

mu is less significant than other parameters, let's try without mu: *(just for fun)*

### GARCH(1,1) - without mu

```{r}
k.garch11a <- garchFit(~garch(1, 1),
                      data = portfolio$PORTFOLIO_r,
                      include.mean = FALSE,
                      cond.dist = "norm", 
                      trace = FALSE) 
summary(k.garch11a)

plot(k.garch11a, which = 10)
plot(k.garch11a, which = 11)
```

-   All parameters are significant.
-   The Ljung-Box test for R\^2 does not show autocorrelation at 5% level between the current and past standardized squared residuals (variance).
-   ACF for R\^2 looks much better.

### AR(5)-GARCH(1,1)

```{r}
k.ar5garch11 <- garchFit(~arma(5, 0) + garch(1, 1),
                         data = portfolio$PORTFOLIO_r,
                         include.mean = TRUE,
                         cond.dist = "norm",
                         trace = FALSE)
summary(k.ar5garch11)
plot(k.ar5garch11, which = 10)
plot(k.ar5garch11, which = 11)
```

-   Except AR2,AR3,AR4,AR5, other parameters are significant.
-   The Ljung-box test for R\^2 shows there is no more autocorrelation between the current and past standardized squared residuals. Similarly, there is no autocorrelation among returns after adding AR part now.
-   Both ACF figures show no significant lags.

Since moast of AR parameters are insignificant, let's try simplier version:

### AR(1)-GARCH(1,1)

```{r}
k.ar1garch11 <- garchFit(~arma(1, 0) + garch(1, 1),
                         data = portfolio$PORTFOLIO_r,
                         include.mean = TRUE,
                         cond.dist = "norm",
                         trace = FALSE)
summary(k.ar1garch11)
plot(k.ar1garch11, which = 10)
plot(k.ar1garch11, which = 11)
```

-   All parameters are significant.
-   The Ljung-box test for R\^2 shows there is no more autocorrelation between the current and past standardized squared residuals. Similarly, there is no autocorrelation among returns after adding AR part now.
-   Both ACF figures show no significant lags.

### AR(1)-ARCH(10)

```{r}
tic()
k.ar1arch10 <- garchFit(~arma(1, 0) + garch(10, 0),
                        data = portfolio$PORTFOLIO_r,
                        include.mean = TRUE,
                        cond.dist = "norm",
                        trace = FALSE)
toc()

summary(k.ar1arch10)
plot(k.ar1arch10, which = 10)
plot(k.ar1arch10, which = 11)
```

-   All parameters are significant.
-   The Ljung-box test for R\^2 shows there is no more autocorrelation between the current and past standardized squared residuals. Similarly, there is no autocorrelation among returns after adding AR part now.
-   Both ACF figures show no significant lags.

### Result: ARCH vs GARCH

Let's compare the models using above mentioned fucntion:

```{r}
compare_ICs_GARCH(c("k.arch1", 
                    "k.arch1a", 
                    "k.arch5", 
                    "k.arch10", 
                    "k.garch11", 
                    "k.garch11a",
                    "k.ar5garch11",
                    "k.ar1garch11", 
                    "k.ar1arch10"))
```

The comparison shows that 8th - **AR(1)-GARCH(1,1)** is the best model among others.

I will use this model for other GARCH extensions.

Let's turn the page for new chapter! ^.^

# GARCH extensions

## EGARCH model

Let's define a model specification:

### AR(1)-EGARCH(1,1)

```{r}
spec <- ugarchspec(# variance equation
    variance.model = list(model = "eGARCH", 
                          garchOrder = c(1, 1)),
    mean.model = list(armaOrder = c(1, 0), 
                      include.mean = TRUE), 
    # assumed distribution of errors
    distribution.model = "norm")
```

and estimate the model:

```{r}
k.ar1egarch11 <- ugarchfit(spec = spec, 
                           data = portfolio$PORTFOLIO_r)

#Check the results:

k.ar1egarch11
```

-   Except mu, other parameters are significant.

-   alpha1 is the asymmetry term of EGARCH model (measuring the leverage effect) and it is negative. So the asymmetry is found.

-   No further autocorrelation in R and R\^2.

-   The Nyblom stability test verifies the null that all parameters or each parameter separately are stable over time (no structural breaks in the model). If test statistic is smaller than the critical value, the null isn't rejected.

    So here we cannot reject the stability jointly (1.1442 \< 1.68), and for any parameter (stat \< 0.47).

-   The Sign Bias Test verifies the null hypothesis if there is still some asymmetric reaction of variance of residuals to shocks. If null not rejected - the asymmetry has been captured correctly in the estimated model. The model does not require higher order.

-   The Adjusted Pearson Goodness-of-Fit test statistics assess the adequacy of the GARCH model in capturing the characteristics of the data. With p_value \< 0.05, we reject the null hypothesis.

```{r}

plot(k.ar1egarch11, which = 3)
plot(k.ar1egarch11, which = 11)
plot(k.ar1egarch11, which = 12)
```

-   The News Impact curve shows the asymmetric effect of shocks, since there is no effect when ϵt−1 is positive but a large effect when the shock is negative.

### AR(1)-EGARCH(1,1) without mu

```{r}
speca <- ugarchspec(# variance equation
  variance.model = list(model = "eGARCH", 
                        garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = FALSE), 
  # assumed distribution of errors
  distribution.model = "norm")

# and estimate the model:

k.ar1egarch11a <- ugarchfit(spec = speca, 
                           data = portfolio$PORTFOLIO_r)
```

Result:

```{r}
k.ar1egarch11a
```

```{r}
plot(k.ar1egarch11a, which = 3) 
plot(k.ar1egarch11a, which = 11) 
plot(k.ar1egarch11a, which = 12)
```

-   Almost the same result with the previous model, just without mu

## The GARCH-t model

### AR(1) - GARCH-st(1,1)

Let's first define a model specification:

```{r}
spec <- ugarchspec(# variance equation
  variance.model = list(model = "sGARCH", 
                        garchOrder = c(1, 1)),
  # mean equation
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE), 
  # assumed distribution of errors
  distribution.model = "std") # std = t-Student


# Then, estimate the model:

k.ar1garcht11 <- ugarchfit(spec = spec, 
                           data = portfolio$PORTFOLIO_r)
# Result:
k.ar1garcht11
```

-   The shape paramters denotes the number of degrees of freedom for the t-Student distribution. In this case it equals to 6.011

-   According to Ljung-Box Test, there is no autocorellation, and according to ARCH LM test, there is no ARCH effect.

-   Nyblom stability test: if test statistic is greater than the critical value the null is rejected. The null hypothesis is about that all parameters or each parameter separately are stable over time (no structural breaks in the model). So here we reject the stability jointly (27 \> 1.68), but cannot reject for any parameter (stat \< 0.47), except omega.

-   The Sign Bias Test verifies the null hypothesis if there is still some asymmetric reaction of variance of residuals to shocks. So the null is rejected the asymmetry has not been captured correctly in the estimated model.

-   The Adjusted Pearson Goodness-of-Fit test statistics assess the adequacy of the GARCH model in capturing the characteristics of the data.

```{r}

plot(k.ar1garcht11, which = 3)
plot(k.ar1garcht11, which = 11)
plot(k.ar1garcht11, which = 12)
```

## The GARCH-in-Mean model

### AR(1)-GARCH-m(1,1)

Let's first define a model specification:

```{r}
spec <- ugarchspec(# variance equation
  variance.model = list(model = "sGARCH", 
                        # sGARCH = standard GARCH
                        garchOrder = c(1, 1)),
  # mean equation - lets turn on the intercept term
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE,
                    # we add an element to the mean equation,
                    # which can be either stdev (archpow 1)
                    # or var (archpow=2)
                    archm = TRUE, archpow = 1), 
  # assumed distribution of errors
  distribution.model = "norm")

# Then, we can estimate the model:

k.ar1garchm11 <- ugarchfit(spec = spec, 
                           data = portfolio$PORTFOLIO_r)

# Let's examine the results:

k.ar1garchm11
```

```{r}
plot(k.ar1garchm11, which = 3)
plot(k.ar1garchm11, which = 11)
plot(k.ar1garchm11, which = 12)
```

**mu** is insignificant, let's remove mu.

### AR(1)-GARCH-m(1,1) without mu

```{r}
spec <- ugarchspec(# variance equation
  variance.model = list(model = "sGARCH", 
                        # sGARCH = standard GARCH
                        garchOrder = c(1, 1)),
  # mean equation - lets turn on the intercept term
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = FALSE,
                    # we add an element to the mean equation,
                    # which can be either stdev (archpow 1)
                    # or var (archpow=2)
                    archm = TRUE, archpow = 1), 
  # assumed distribution of errors
  distribution.model = "norm")

# Model estimation:

k.ar1garchm11a <- ugarchfit(spec = spec, 
                           data = portfolio$PORTFOLIO_r)

# Model summary:

k.ar1garchm11a
```

-   Except omega, all parameters are significant.

-   There is no ARCH effect.

-   Ljung-Box Test shows no autocorrelation.

-   The Nyblom stability test verifies the null that all parameters or each parameter separately are stable over time (no structural breaks in the model). If test statistic is smaller than the critical value, the null isn't rejected.

So here we cannot reject the stability jointly (1.093 < 1.68 1.68), and for any parameter (stat < 0.47).

-   The Sign Bias Test verifies the null hypothesis if there is still some asymmetric reaction of variance of residuals to shocks. So the null is rejected, the asymmetry has not been captured correctly in the estimated model.

-   The Adjusted Pearson Goodness-of-Fit test statistics assess the adequacy of the GARCH model in capturing the characteristics of the data. With p_value < 0.05, we reject the null hypothesis.

```{r}

plot(k.ar1garchm11a, which = 3)
plot(k.ar1garchm11a, which = 11)
plot(k.ar1garchm11a, which = 12)
```

## The EGARCH in mean-t model

### AR(1)-EGARCH-m-st(1,1)

Let's first define the model specification:

```{r}
spec <- ugarchspec(# variance equation
  variance.model = list(model = "eGARCH", 
                        garchOrder = c(1, 1)),
  # mean equation
  mean.model = list(armaOrder = c(1, 0), 
                    include.mean = TRUE,
                    archm = TRUE, archpow = 1), 
  # assumed distribution of errors
  distribution.model = "std") # std = t-Student

# Now, let's estimate the model:

k.ar1egarchmt11 <- ugarchfit(spec = spec, 
                             data = portfolio$PORTFOLIO_r)

# The model summary:

k.ar1egarchmt11
```
* All parameteres are significant.
* There is no Autocorellation - Ljung-Box Test
* No Arch effect
* Nyblom stability test 1.86 < 2.11
* The Sign Bias Test verifies the null hypothesis if there is still some asymmetric reaction of variance of residuals to shocks. If null not rejected - the asymmetry has been captured correctly in the estimated model. The model does not require higher order.
* The Adjusted Pearson Goodness-of-Fit test statistics assess the adequacy of the GARCH model in capturing the characteristics of the data. With p_value > 0.05, we do not reject the null hypothesis.

```{r}
plot(k.ar1egarchmt11, which = 3)
plot(k.ar1egarchmt11, which = 11)
plot(k.ar1egarchmt11, which = 12)
```

# Comparison

```{r}
compare_ICs_ugarchfit(c("k.ar1egarch11",
                        "k.ar1egarch11a",
                        "k.ar1garcht11",
                        "k.ar1garchm11", 
                        "k.ar1garchm11a",
                        "k.ar1egarchmt11"))
```

According to the comparison, last model, ***AR(1)-EGARCH-m-st(1,1)*** is the best model.

Let's turn page to the new chapter 🍀

# VaR (Value at risk)
